---
title: "Final Analyses"
output:
  html_document: 
    keep_md: yes
---

## Creating the Dataset: 

```{r}
library(ggplot2)
library(car)
library(lme4)
library(lmerTest)
library(plotrix)
library(Rmisc)
library(dplyr)
library(MASS)
library(MuMIn)
library(ggthemes)
```

### Calculations for additional metrics:

```{r}
thesis.data <- read.table("thesis_data_June10.csv", header = T, sep = ",")

thesis.data$survival <- (thesis.data$count.after / thesis.data$count.before)
thesis.data$midden <- as.factor(thesis.data$midden)
thesis.data$lw.ratio <- thesis.data$length / thesis.data$width 
thesis.data$relative.fitness <- thesis.data$survival / mean(thesis.data$survival) 
thesis.data$mass.indiv.seed <- thesis.data$mass.seed / thesis.data$count.seeds
thesis.data$mass.ratio <- thesis.data$mass.seed / thesis.data$mass.cone
thesis.data$wl.ratio <- thesis.data$width / thesis.data$length

thesis.data <- thesis.data[thesis.data$id != "2.12.15", ]

head(thesis.data)
```

### Standardizing the cone traits:

```{r}
z.thesis.data <- thesis.data

z.thesis.data$length <- scale(z.thesis.data$length, center = TRUE, scale = TRUE)
z.thesis.data$width <- scale(z.thesis.data$width, center = TRUE, scale = TRUE)
z.thesis.data$width.scale <- scale(z.thesis.data$width.scale, center = TRUE, scale = TRUE)
z.thesis.data$count.seeds <- scale(z.thesis.data$count.seeds, center = TRUE, scale = TRUE)
z.thesis.data$mass.ratio <- scale(z.thesis.data$mass.ratio, center = TRUE, scale = TRUE)
z.thesis.data$lw.ratio <- scale(z.thesis.data$lw.ratio, center = TRUE, scale = TRUE)
z.thesis.data$mass.indiv.seed <- scale(z.thesis.data$mass.indiv.seed, center = TRUE, scale = TRUE)
z.thesis.data$wl.ratio <- scale(z.thesis.data$wl.ratio, center = TRUE, scale = TRUE)

head(z.thesis.data)
```

### Splitting into NS and S Cohorts:

```{r}
z.thesis.data.NS <- z.thesis.data[z.thesis.data$serotiny == "NS",]
z.thesis.data.S <- z.thesis.data[z.thesis.data$serotiny == "S",]
```


## Table of values:
  - means of each trait (for serotinous and non-serotinous)
  - SEs
  - ranges

### Means:  
```{r}
thesis.subset <- subset(thesis.data, select = c(distance, serotiny, cones.sampled, length, width, width.scale, count.seeds, mass.cone, mass.seed, mass.ratio, count.before, count.after, survival, lw.ratio))

#Had to figure out a way to not get undefined for the ones with no seeds

thesis.subset$mass.indiv.seed <- thesis.subset$mass.seed / (thesis.subset$count.seeds + .000001)

#avg.mean <- aggregate(thesis.subset, list(serotiny = thesis.subset$serotiny), FUN = mean)
```

### Individual SE calculations per trait:  
```{r}
SE.distance <- summarySE(thesis.subset, measurevar = "distance", groupvars = "serotiny")
SE.cones.sampled <- summarySE(thesis.subset, measurevar = "cones.sampled", groupvars = "serotiny")
SE.length <- summarySE(thesis.subset, measurevar = "length", groupvars = "serotiny")
SE.width<- summarySE(thesis.subset, measurevar = "width", groupvars = "serotiny")
SE.width.scale <- summarySE(thesis.subset, measurevar = "width.scale", groupvars = "serotiny")
SE.count.seeds <- summarySE(thesis.subset, measurevar = "count.seeds", groupvars = "serotiny")
SE.mass.cone <- summarySE(thesis.subset, measurevar = "mass.cone", groupvars = "serotiny")
SE.mass.indiv.seed <- summarySE(thesis.subset, measurevar = "mass.indiv.seed", groupvars = "serotiny")
SE.mass.ratio <- summarySE(thesis.subset, measurevar = "mass.ratio", groupvars = "serotiny")
SE.count.before <- summarySE(thesis.subset, measurevar = "count.before", groupvars = "serotiny")
SE.count.after <- summarySE(thesis.subset, measurevar = "count.after", groupvars = "serotiny")
SE.survival <- summarySE(thesis.subset, measurevar = "survival", groupvars = "serotiny")
SE.lw.ratio <- summarySE(thesis.subset, measurevar = "lw.ratio", groupvars = "serotiny")

```

### T-tests for trait data:
```{r}
thesis.subset.NS <- thesis.subset[thesis.data$serotiny == "NS",]
thesis.subset.S <- thesis.subset[thesis.data$serotiny == "S",]
```

```{r}
t.test(thesis.subset.NS$distance, thesis.subset.S$distance)
t.test(thesis.subset.NS$cones.sampled, thesis.subset.S$cones.sampled)
t.test(thesis.subset.NS$length, thesis.subset.S$length)
t.test(thesis.subset.NS$width, thesis.subset.S$width)
t.test(thesis.subset.NS$width.scale, thesis.subset.S$width.scale)
t.test(thesis.subset.NS$count.seeds, thesis.subset.S$count.seeds)
t.test(thesis.subset.NS$mass.cone, thesis.subset.S$mass.cone)
t.test(thesis.subset.NS$mass.indiv.seed, thesis.subset.S$mass.indiv.seed)
t.test(thesis.subset.NS$mass.ratio, thesis.subset.S$mass.ratio)
t.test(thesis.subset.NS$count.before, thesis.subset.S$count.before)
t.test(thesis.subset.NS$count.after, thesis.subset.S$count.after)
t.test(thesis.subset.NS$lw.ratio, thesis.subset.S$lw.ratio)
```


### T-test for survival
```{r}
t.test(z.thesis.data.NS$survival, z.thesis.data.S$survival)
```

## Selection Gradients: 

### 3 cone traits of interest
```{r}
additive <- glm(relative.fitness ~ lw.ratio + mass.ratio + serotiny + width.scale, data = z.thesis.data)

summary(additive)

```

### Cone traits of interest and potential confounding variables

**Utilized gradient analysis** 
```{r}
alltraits <- glm(relative.fitness ~ lw.ratio + mass.ratio + serotiny + width.scale + jdate.cone + dbh + distance + direction + midden.S, data = z.thesis.data)

summary(alltraits)
```

### Double-checking Midden Serotiny level 

```{r}
checkingmiddenS <- glm(relative.fitness ~ lw.ratio + mass.ratio + width.scale + serotiny + midden.S, data = z.thesis.data)

summary(checkingmiddenS)
```

Nope, not significant!! 

### Checking for interactions: 

```{r}
checkingmiddenSint <- glm(relative.fitness ~ mass.ratio*serotiny*midden.S, data = z.thesis.data)

summary(checkingmiddenSint)
```

### Interactions with traits of interest:
```{r}
allinteractions <- glm(relative.fitness ~ mass.ratio*serotiny*lw.ratio*width.scale, data = z.thesis.data)
summary(allinteractions)
```

### Significant terms from that model:
```{r}
sigfromall <- glm(relative.fitness ~ mass.ratio + serotiny + mass.ratio:serotiny:lw.ratio, data = z.thesis.data)

summary(sigfromall)
```

### Simple interactive:
```{r}
simpleinteractive <- glm(relative.fitness ~ mass.ratio*serotiny, data = z.thesis.data)

summary(simpleinteractive)
```

### Simple additive:
```{r}
simpleadditive <- glm(relative.fitness ~ mass.ratio + serotiny, data = z.thesis.data)

summary(simpleadditive)

```


### AICc Calculations
```{r}
AICvals <- AIC(additive, alltraits, checkingmiddenS, checkingmiddenSint, allinteractions, sigfromall, simpleinteractive, simpleadditive)
AICvals$corrected <- c(AICc(additive), AICc(alltraits), AICc(checkingmiddenS), AICc(checkingmiddenSint), AICc(allinteractions), AICc(sigfromall), AICc(simpleinteractive), AICc(simpleadditive))
AICvals$deltacorr <- AICvals$corrected - min(AICvals$corrected)

AICvals
```

## VIF Calculations:
```{r}
#In Table Order:
vif(alltraits) #NO
vif(allinteractions) #YES - interactions, duh 
vif(additive) #NO
vif(simpleinteractive) #YES - interaction again
vif(simpleadditive) #NO


vif(glm(relative.fitness ~ length + width + mass.cone + mass.seed + serotiny + width.scale + jdate.cone + dbh + distance + direction + midden.S, data = z.thesis.data)) #YES
vif(glm(relative.fitness ~ length + width + mass.cone + width.scale, data = z.thesis.data)) #YES
vif(glm(relative.fitness ~ length + width, data = z.thesis.data)) #YES
vif(glm(relative.fitness ~ length + mass.cone, data = z.thesis.data)) #NO
vif(glm(relative.fitness ~ length + width.scale, data = z.thesis.data)) #NO
vif(glm(relative.fitness ~ width + mass.cone, data = z.thesis.data)) #YES
vif(glm(relative.fitness ~ width + width.scale, data = z.thesis.data)) #YES (barely)
vif(glm(relative.fitness ~ mass.cone + width.scale, data = z.thesis.data)) #NO

vif(glm(relative.fitness ~ lw.ratio + width.scale, data = z.thesis.data)) #NO
vif(glm(relative.fitness ~ lw.ratio + mass.cone, data = z.thesis.data)) #NO

```

## Selection Differentials:

### Length to Width Ratio
```{r}
lwmodNS <- glm(relative.fitness ~ lw.ratio, data = z.thesis.data.NS)

summary(lwmodNS)
```

```{r}
lwmodS <- glm(relative.fitness ~ lw.ratio, data = z.thesis.data.S)

summary(lwmodS)
```

### Cone to Seed Mass Ratio
```{r}
massmodNS <- glm(relative.fitness ~ mass.ratio, data = z.thesis.data.NS)

summary(massmodNS)
```

```{r}
massmodS <- glm(relative.fitness ~ mass.ratio, data = z.thesis.data.S)

summary(massmodS)
```

### Scale Width
```{r}
widthmodNS <- glm(relative.fitness ~ width.scale, data = z.thesis.data.NS)

summary(widthmodNS)
```

```{r}
widthmodS <- glm(relative.fitness ~ width.scale, data = z.thesis.data.S)

summary(widthmodS)
```

### Number of Full Seeds

```{r}
numseedNS <- glm(relative.fitness ~ count.seeds, data = z.thesis.data.NS)

summary(numseedNS)
```

```{r}
numseedS <- glm(relative.fitness ~ count.seeds, data = z.thesis.data.S)

summary(numseedS)
```

### Individual Seed Mass:

```{r}
seedmassNS <- glm(relative.fitness ~ mass.indiv.seed, data = z.thesis.data.NS)

summary(seedmassNS)
```

```{r}
seedmassS <- glm(relative.fitness ~ mass.indiv.seed, data = z.thesis.data.S)

summary(seedmassS)
```

### Cone Length

```{r}
conelengthNS <- glm(relative.fitness ~ length, data = z.thesis.data.NS)

summary(conelengthNS)
```

```{r}
conelengthS <- glm(relative.fitness ~ length, data = z.thesis.data.S)

summary(conelengthS)
```

### Cone Width

```{r}
conewidthNS <- glm(relative.fitness ~ width, data = z.thesis.data.NS)

summary(conewidthNS)
```

```{r}
conewidthS <- glm(relative.fitness ~ width, data = z.thesis.data.S)

summary(conewidthS)
```

### Cone Mass

```{r}
conemassNS <- glm(relative.fitness ~ mass.cone, data = z.thesis.data.NS)

summary(conemassNS)
```

```{r}
conemassS <- glm(relative.fitness ~ mass.cone, data = z.thesis.data.S)

summary(conemassS)
```

### Midden Serotiny

```{r}
middenNS <- glm(relative.fitness ~ midden.S, data = z.thesis.data.NS)

summary(middenNS)
```

```{r}
middenS <- glm(relative.fitness ~ midden.S, data = z.thesis.data.S)

summary(middenS)
```

```{r}
midden <- glm(relative.fitness ~ midden.S, data = z.thesis.data)

summary(midden)
```

## Graphs

### Graph of Relative Fitness vs. Serotiny 
```{r}
ggplot(z.thesis.data, aes(x=serotiny, y=survival)) + 
  geom_boxplot(aes()) +
  labs(x = "", y = "Cone Survival Rate") + 
  theme_bw() + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 12)) +
  scale_x_discrete(labels = c("Non-serotinous trees", "Serotinous trees"))

ggsave("serotiny.boxplot.14Dec.png", width = 8, height = 5)
```

### Graph of Relative Fitness vs. Kernel to Cone Mass Ratio

```{r}
ggplot(z.thesis.data, aes(x=mass.ratio, y=relative.fitness, shape = serotiny)) +
  theme_bw() + 
  geom_point(aes(), size = 2) +
  scale_shape_manual(values = c(1,19), 
                     labels = c("Non-serotinous", "Serotinous")) + 
  theme(legend.position = c(.85,.9)) + 
  geom_smooth(method = "lm", aes(linetype = serotiny), color = "black") +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("Non-serotinous", "Serotinous")) +
  theme(legend.position = c(.85, .9)) + 
  theme(legend.text = element_text(size = 12)) + 
  theme(axis.line.x = element_line(color="black"),
        axis.line.y = element_line(color="black")) + 
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16)) + 
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)) + 
  labs(x = "Standardized Kernel to Cone Mass Ratio", y = "Relative Fitness") + 
  theme(legend.title = element_blank()) 
  
  ggsave("serotiny.massratio.14Dec.png", width = 8, height = 5)
```

### Graph of Relative Fitness vs. Individual Kernel Mass

```{r}
ggplot(z.thesis.data, aes(x=mass.indiv.seed, y=relative.fitness, shape = serotiny)) +
  theme_bw() + 
  geom_point(aes(), size = 2) +
  scale_shape_manual(values = c(1,19), 
                     labels = c("Non-serotinous", "Serotinous")) + 
  theme(legend.position = c(.85,.9)) + 
  geom_smooth(method = "lm", aes(linetype = serotiny), color = "black") +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("Non-serotinous", "Serotinous")) +
  theme(legend.position = c(.85, .9)) + 
  theme(legend.text = element_text(size = 12)) + 
  theme(axis.line.x = element_line(color="black"),
        axis.line.y = element_line(color="black")) + 
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16)) + 
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)) + 
  labs(x = "Standardized Individual Kernel Mass", y = "Relative Fitness") + 
  theme(legend.title = element_blank()) 
  
  ggsave("serotiny.kernelmass.15Dec.png", width = 8, height = 5)
```

### With that outlier removed: 

```{r}
ind.thesis.data <- z.thesis.data[z.thesis.data$id != "2.04.17", ]

ggplot(ind.thesis.data, aes(x=mass.indiv.seed, y=relative.fitness)) + 
  geom_point(aes(color = serotiny)) +
  geom_smooth(method = "lm", aes(color = serotiny)) +
  labs(x = "Individual Seed Mass", y = "Relative Fitness")
```

### Graph of Relative Fitness vs. Number of Kernels 

```{r}
ggplot(z.thesis.data, aes(x=count.seeds, y=relative.fitness, shape = serotiny)) +
  theme_bw() + 
  geom_point(aes(), size = 2) +
  scale_shape_manual(values = c(1,19), 
                     labels = c("Non-serotinous", "Serotinous")) + 
  theme(legend.position = c(.85,.9)) + 
  geom_smooth(method = "lm", aes(linetype = serotiny), color = "black") +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("Non-serotinous", "Serotinous")) +
  theme(legend.position = c(.85, .9)) + 
  theme(legend.text = element_text(size = 12)) + 
  theme(axis.line.x = element_line(color="black"),
        axis.line.y = element_line(color="black")) + 
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16)) + 
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)) + 
  labs(x = "Standardized Number of Kernels per Cone", y = "Relative Fitness") + 
  theme(legend.title = element_blank()) 
  
  ggsave("serotiny.kernelcount.15Dec.png", width = 8, height = 5)
```

